Effectful Semantics for Substructural SSA

### 2023-11-13 pldi24/paper.tex
Tap

\TODO{text, \cite{sparky}}

\begin{definition}[Pomset] 
  A \textbf{pomset} over a set of actions \(\mc{A}\) with a distinguished null action "\textbf{tick}" \(\delta \in \mc{A}\)
  \TODO{finish, clean}
\end{definition}

\TODO{Form a monoid, so we can define...}

Let \(\ms{Pom}\) denote the monoid of pomsets...

\TODO{stream action... trace monad \(\ms{Traces}\;\ms{Pom}\;\ms{Pom}\)...}

\TODO{The SC Monad, which, via composition w/ the state transformer w/ a buffer as state, lets us}

\TODO{Build The Weak Memory Monad, which is too powerful, so we do...}

\begin{definition}[Monadic Predicate] \
  \TODO{this and naming}
\end{definition}

\TODO{Proper Weak Memory Monad, preserves the Elgot structure (yey)}

>>>>

Another interesting use of the trace monad to build a sophisticated \isotopessa model is to build a model of TSO-style weak memory based on \citet{sparky}. We begin with the following definitions:
\begin{definition}[Pomset] 
  A \textbf{pomset} \(\alpha\) over a set of actions \(\mc{A}\) with a
  distinguished null action "\textbf{tick}" \(\delta \in \mc{A}\) is a
  partially-ordered \textbf{carrier set} \(P\) such that every \(p \in P\) has
  finitely many predecessors equipped with a mapping \(\alpha: P \to \mc{A}\)
  quotiented under the equivalence relation \(\alpha: P \to \mc{A} \simeq
  \alpha': Q \to \mc{A}\) if, after removing finitely many \(\delta\) from \(P\)
  and \(Q\), there exists a partial order isomorphism between them which
  respects the values of \(\alpha, \alpha'\). A pomset is \textbf{finite} if its
  carrier set is.
\end{definition}
(Finite) pomsets form a monoid under sequential composition \(\alpha;\beta\), which is defined by the function \([\alpha, \beta]: \ms{trim}(P + Q) \to \mc{A}\), where \(P + Q\) is given the lexicographic ordering and \(\ms{trim}(P)\) removes all elements of \(P\) with infinitely many predecessors. They also form a monoid under parallel composition \(\alpha || \beta\), which is defined by the function \([\alpha, \beta]: P + Q \to \mc{A}\) where \(P + Q\) is given the standard partial ordering (with elements of \(P\) and \(Q\) incomparable).

Given a partially ordered set \(N\) and a family of pomsets \(\alpha_n\) for \(n \in N\), we can define their \textit{sum} \(\Sigma_n\alpha_n\) to be given by the function \((n, a) \mapsto \alpha_n\;a: \ms{trim}(\Sigma_nP_n) \to \mc{A}\), where the dependent product \(\Sigma_nP_n\) is given the lexicographic order. In particular, choosing \(N = \nats\) makes \(\Sigma: \ms{Pom}_{\ms{fin}}^\omega \to \ms{Pom}\) into a stream action w.r.t the sequential composition monoid on finite pomsets, which allows us to define the \textbf{sequential consistency monad} \(\ms{SC}\;A = \ms{Traces}\;\Sigma\;A\).

\TODO{how to do reads}

\TODO{how to do writes}

\TODO{weak memory monad}

\TODO{new read/write}

\TODO{fence}

\TODO{submonad rune}

\TODO{Buffer continuity}
